Journal and article coverage

### 2025-02-05 analysis/manuscript.Rmd
Describe data sources used

The aim of this study is to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
To achieve this, a replication study is conducted by comparing results  from hoaddata, an openly available dataset on hybrid open access uptake, with the established bibliometric databases Web of Science and Scopus.
Replication studies, which can be found in the Life, Natural and Social Sciences, typically employ the same or comparable methods as the original study to determine whether consistent results can be obtained (https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/).

This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network for Bibliometrics (Kompetenznetzwerk Bibliometrie, KB), which comprises Web of Science and Scopus snapshots.
After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using author roles (first and corresponding) and affiliation data are described.
Overall, xxxx hybrid journals from xxx agreements formed the basis of this study. This study covered journal articles published between 2019 and 2023.

## Dataset characteristics

### {hoaddata}

@@

@@

@@

```{r dataset_filtering, cache=TRUE}
## HOAD
hoad_jns <- hoaddata::jct_hybrid_jns |>
  filter(issn_l != "0027-8424")

# Full OA journals detected through OA proportion > .95
hoad_oa_excluded <- bq_table_download("hoa-article.hoaddata_sep24.cc_oa_prop") |>
  distinct(issn_l)

active_jns <- hoad_jns |>
  # Only active hybrid journals from JCT
  inner_join(hoaddata::jn_ind, by = "issn_l") |>
  filter(!cr_year %in% c("2017", "2018", "2024"))

active_jns_with_oa <- active_jns |> 
  filter(!is.na(cc))

# Get indicators for all articles without applying paratext adn supplement detection
hoad_all <- read_csv(here::here("data", "cr_all_ind.csv"))

## Web of Science
wos_jct_items_df <- readr::read_csv(here::here("data-raw", "wos_jct_items_df.csv"))

wos_hybrid_jns <- wos_jct_items_df |>
  # No PNAS
  filter(issn_l != "0027-8424") |>
  # No Full OA detected by OA prop
  filter(!issn_l %in% hoad_oa_excluded$issn_l)

wos_active_jns <- wos_hybrid_jns |>
  # Use online first, if available
  mutate(online_year = lubridate::year(wos_pubdate_online)) |>
  mutate(pub_year = ifelse(!is.na(online_year), online_year, pubyear)) |>
  filter(pub_year %in% 2019:2023)

wos_active_jns_with_oa <- wos_active_jns |> 
  filter(grepl("hybrid", oa_status))

# wos core 
wos_active_core <- wos_active_jns |>
  filter(grepl("Article|^Review$", item_type))

wos_active_core_oa <- wos_active_jns |>
   filter(grepl("hybrid", oa_status))

## Scopus

scp_jct_items_df <- readr::read_csv(here::here("data-raw", "scp_jct_items_df.csv"))

scp_hybrid_jns <- scp_jct_items_df |>
  # No PNAS
  filter(issn_l != "0027-8424") |>
  # No Full OA detected by OA prop
  filter(!issn_l %in% hoad_oa_excluded$issn_l)

scp_active_jns <- scp_hybrid_jns |>
   # Use online first, if available
  mutate(pub_year = ifelse(!is.na(first_pubyear), first_pubyear, pubyear)) |>
  filter(pub_year %in% 2019:2023)

scp_active_jns_with_oa <- scp_active_jns |> 
  filter(grepl("hybrid", oa_status))

# core
scp_active_core <- scp_active_jns |>
   filter(grepl("^Article|^Review", item_type))

scp_active_core_oa <- scp_active_core |>
    filter(grepl("hybrid", oa_status))

```

```{r journals_overview_table, cache=TRUE}
### Journals
ind_active_jns <- c(`hoad` = hoad_all$active_journals, 
                    `wos` = length(unique(wos_active_jns$issn_l)),
                    `scp` = length(unique(scp_active_jns$issn_l)))
ind_active_jns_core <- c(`hoad` = length(unique(active_jns$issn_l)), 
                    `wos` = length(unique(wos_active_core$issn_l)),
                    `scp` = length(unique(scp_active_core$issn_l)))
ind_active_jns_core_with_oa <-  c(`hoad` = length(unique(active_jns_with_oa$issn_l)), 
                    `wos` = length(unique(wos_active_core_oa$issn_l)),
                    `scp` = length(unique(scp_active_core_oa$issn_l)))

jns_ind <- bind_rows(ind_active_jns = ind_active_jns,
          ind_active_jns_core = ind_active_jns_core,
          ind_active_jns_core_with_oa = ind_active_jns_core_with_oa, .id = "indicator")
```

```{r article_overview_table, cache=TRUE}
### Article Volume

## hoad

# DOI set too large to track it with Git, see 001_data_gathering how the data
# was obtained from Google BQ
cr_df <- readr::read_csv(here::here("data-raw/hoad_dois_all_19_23.csv"))

ind_articles_hoad <- cr_df |>
  filter(issn_l %in% active_jns_with_oa$issn_l) |>
  distinct(doi) |>
  nrow()

## WoS
ind_articles_wos <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  distinct(item_id) |>
  nrow()

# core
ind_core_articles_wos <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  distinct(item_id) |>
  nrow()
# doi
ind_articles_wos_doi <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  distinct(doi) |>
  nrow()

ind_core_articles_wos_doi <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  distinct(doi) |>
  nrow()
## scopus
ind_articles_scp <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  distinct(item_id) |>
  nrow()
# core
ind_core_articles_scp <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  filter(grepl("^Article|^Review", item_type)) |>
  distinct(item_id) |>
  nrow()
# doi
ind_articles_scp_doi <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  distinct(doi) |>
  nrow()

ind_core_articles_scp_doi <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  distinct(doi) |>
  nrow()

# Bringing it all together
ind_articles_total <- c(`hoad` = hoad_all$active_journals_all,
                        `wos` = ind_articles_wos,
                        `scp` = ind_articles_scp)
ind_articles_core <- c(`hoad` = ind_articles_hoad,
                        `wos` = ind_core_articles_wos,
                        `scp` = ind_core_articles_scp)
ind_articles_doi <- c(`hoad` = hoad_all$active_journals_all,
                        `wos` = ind_articles_wos_doi,
                        `scp` = ind_articles_scp_doi)
ind_articles_core_doi <- c(`hoad` = ind_articles_hoad,
                        `wos` = ind_core_articles_wos_doi,
                        `scp` = ind_core_articles_scp_doi)

articles_ind <- bind_rows(ind_articles_total = ind_articles_total,
                          ind_articles_core = ind_articles_core,
                          ind_articles_doi = ind_articles_doi,
                          ind_articles_core_doi = ind_articles_core_doi, .id = "indicator" )

## oa just core with doi

### hoad
ind_oa_articles_hoad <- hoaddata::cc_articles |>
  filter(between(cr_year, 2019, 2023)) |>
  filter(issn_l %in% active_jns$issn_l) |>
  distinct(doi) |>
  nrow()

### wos

ind_oa_articles_wos_all <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("hybrid", oa_status)) |>
  distinct(item_id) |>
  nrow()

ind_oa_articles_wos_core <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  filter(grepl("hybrid", oa_status)) |>
  distinct(item_id) |>
  nrow()

### scp
ind_oa_articles_scp_all <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
   filter(grepl("hybrid", oa_status)) |>
  distinct(item_id) |>
  nrow()
ind_oa_articles_scp_core <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  filter(grepl("hybrid", oa_status)) |>
  distinct(item_id) |>
  nrow()

ind_oa_articles_all <- c(`hoad` = hoad_all$ind_oa_articles_all,
                                             `wos` = ind_oa_articles_wos_all,
                                             `scp` = ind_oa_articles_scp_all
                     )
ind_oa_articles_core <-  c(`hoad` = ind_oa_articles_hoad,
                                             `wos` = ind_oa_articles_wos_core,
                                             `scp` = ind_oa_articles_scp_core
                     )

oa_articles_ind <- bind_rows(ind_oa_articles_all = ind_oa_articles_all,
                          ind_oa_articles_core = ind_oa_articles_core, .id = "indicator")
```

```{r ind_affiliations}

### affiliations

### WOS
wos_active_core_with_oa_item_ids <-  wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  distinct(item_id, oa_status)

wos_aff <- readr::read_csv(here::here("data-raw", "wos_jct_affiliations.csv"))

wos_aff_df <- wos_aff |>
  inner_join(wos_active_core_with_oa_item_ids, by = "item_id")

wos_first_au_with_affiliation <- wos_aff_df |>
  filter(author_seq_nr == 1, !is.na(vendor_org_id)) |>
  distinct(item_id) |>
  nrow()

wos_cor_au_with_affiliation <- wos_aff_df |>
  filter(corresponding == TRUE, !is.na(vendor_org_id)) |>
  distinct(item_id) |>
  nrow()

# Scopus
scp_active_core_with_oa_item_ids <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  filter(grepl("^Article|^Review", item_type)) |>
  distinct(item_id, oa_status)

scp_aff <-  readr::read_csv(here::here("data-raw", "scp_jct_affiliations.csv"))

scp_aff_df <- scp_aff |>
  inner_join(scp_active_core_with_oa_item_ids, by = "item_id")

scp_first_au_with_affiliation <- scp_aff_df |>
  filter(author_seq_nr == 1, !is.na(vendor_org_id)) |>
  distinct(item_id) |>
  nrow()

scp_cor_au_with_affiliation <- scp_aff_df |>
  filter(corresponding == TRUE, !is.na(vendor_org_id)) |>
  distinct(item_id) |>
  nrow()

# hoad
aff_oalex_stats <- readr::read_csv(here::here("data", "aff_oalex_stats.csv"))

aff_ind <- bind_rows(
  first_aff = c(hoad = unlist(aff_oalex_stats$first_author_with_ror), 
                wos = wos_first_au_with_affiliation,
                scp = scp_first_au_with_affiliation),
  cor_aff = c(hoad = unlist(aff_oalex_stats$corresponding_authors_inst),
              wos = wos_cor_au_with_affiliation,
              scp = scp_cor_au_with_affiliation), .id = "indicator"
)

```

!<-- -- internal  -- external  - matching table institutions - journal-level data - country affiliation data  -->

```{r methods_overview_table, cache=TRUE}
methods_overview_table_df <- bind_rows(
  # Journals
  jns_ind,
  # Articles
  articles_ind,
  # OA
  oa_articles_ind,
  # Affiliation and Auhtor roles
  aff_ind
) |>
  # Human-readable indicator labels
  mutate(
    indicator = case_when(
      indicator == "ind_active_jns" ~ "Active journals",
      indicator == "ind_active_jns_core" ~ "Active journals (core)",
      indicator == "ind_active_jns_core_with_oa" ~ "Active journals (core) with OA",
      indicator == "ind_articles_total" ~ "Total published articles ",
      indicator == "ind_articles_core" ~ "Core articles",
      indicator == "ind_articles_doi" ~ "Articles with DOI",
      indicator == "ind_articles_core_doi" ~ "Core articles with DOI",
      indicator == "ind_oa_articles_all" ~ "OA articles",
      indicator == "ind_oa_articles_core" ~ "Core OA articles",
      indicator == "first_aff" ~ "First author articles",
      indicator == "cor_aff" ~ "Corresponding author articles",
    )
  )

>>>>

The aim of this study is to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
To achieve this, results from hoaddata, an openly available collection of open research information regarding hybrid open access, was compared with the established bibliometric databases Web of Science and Scopus.
After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using open access evidence, author roles (first and corresponding) and affiliation data are presented.
Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study.

<!-- Replication studies, which can be found in the Life, Natural and Social Sciences, typically employ the same or comparable methods as the original study to determine whether consistent results can be obtained (https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/).

This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network for Bibliometrics (Kompetenznetzwerk Bibliometrie, KB), which comprises Web of Science and Scopus snapshots.
After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using author roles (first and corresponding) and affiliation data are presented.
Overall, xxxx hybrid journals from xxx agreements formed the basis of this study. This study covered journal articles published between 2019 and 2023.   -->

## Data sources

### hoaddata

hoaddata, developed and maintained by the author, is an R data package comprising information about the uptake of hybrid open access since 2017 from several openly available data sources.
It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.

More specifically, hoaddata uses Crossref, a DOI registration agency, for obtaining journal publication volume and open access status through
Creative Commons licence information relative to the published version ("version of record").
Because of limited affiliation metadata in Crossref (https://doi.org/10.31222/osf.io/smxe5), hoaddata sources  first-author affiliations from OpenAlex.
While the country in which a first-author was located was used to aggregate country-level statistics, Research Organization Registry (ROR) identifiers (ROR-ID) were used in conjunction with ESAC registry information on the duration of an agreement was used to estimate whether an article was published under active transformative agreements.
This matching benefited from the availability of the ROR-IDs in both sources.
To improve the matching, JCT data were enriched to include associated institutions, such as unviersity hospitals.

hoaddata follows good practices for computational reproducibility using R.
The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow interfaces with the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.
The package has been regularly updated since 2022 and the version including the computation log is available on GitHub.

hoaddata is used as a data basis of the Hybrid Open Access Dashboard, a data analytics services for  library consortia and publishers to track the uptake of hybrid open access through transformative agreements.
It was also used in bibliometric research (Jahn 2025).

@@

Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020]. 
The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).

The WoS provides important data points for studying spending on open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as Openalex. However, it lacks information about journals and articles under transformative agreements.

For programmatic access to article-level data, the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany is used to access bibliometric data. The KB processes raw XML data provided by Clarivate Analytics, which is provided as an in-house PostgreSQL database under a uniform schema.
To support reproducibility, KB maintains annual snapshots of the database. 
Accordingly, this study used the annual snapshot from April 2024, which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].

@@

Elsevier's Scopus, launched in 2004, is another widely used proprietary bibliometric database for measuring research [@Baas_2020].
Similar to the Web of Science, Scopus is selective with regard to the journals it indexes.
However, its coverage is substantially more extensive than that of the Web of Science Core collection [@Singh_2021; @Visser_2021].
With detailed metadata about article types, open access status information derived from Unpaywall, author roles, and disambiguated affiliations, Scopus also contains important data to assess open access uptake, although no direct information regarding transformative agreements was available at the time of the study.

This study used the Scopus annual snapshot of April 2024 as provided by the KB. The same KB curation effort was applied to the  Scopus raw data as for the Web of Science [@schmidt_2024_13935407].

@@

```{r methods_overview_table, cache=TRUE}
load(here::here("data", "manuscript_data.Rda"))
methods_overview_table_df <- bind_rows(
  # Journals
  jns_ind,
  # Articles
  articles_ind,
  # OA
  oa_articles_ind,
  # Affiliation and Auhtor roles
  aff_ind
) |>
  # Human-readable indicator labels
  mutate(
    indicator = case_when(
      indicator == "ind_active_jns" ~ "Active journals",
      indicator == "ind_active_jns_core" ~ "Active journals (core)",
      indicator == "ind_active_jns_core_with_oa" ~ "Active journals (core) with OA",
      indicator == "ind_articles_total" ~ "Total published articles ",
      indicator == "ind_articles_core" ~ "Core articles",
      indicator == "ind_articles_doi" ~ "Articles with DOI",
      indicator == "ind_articles_core_doi" ~ "Core articles with DOI",
      indicator == "ind_oa_articles_all" ~ "OA articles",
      indicator == "ind_oa_articles_core" ~ "Core OA articles",
      indicator == "first_aff" ~ "First author articles",
      indicator == "cor_aff" ~ "Corresponding author articles",
    )
  )

### 2025-02-05 analysis/manuscript.tex
Describe data sources used

% bibliography

@@

The aim of this study is to demonstrate the suitability of open
scholarly data sources for assessing the impact of transformative
agreements on hybrid open access. To achieve this, a replication study
is conducted by comparing results from hoaddata, an openly available
dataset on hybrid open access uptake, with the established bibliometric
databases Web of Science and Scopus. Replication studies, which can be
found in the Life, Natural and Social Sciences, typically employ the
same or comparable methods as the original study to determine whether
consistent results can be obtained
(\url{https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/}).

This study compares results drawn from the openly available and
regularly updated hoaddata, an R package leveraging the open scholarly
data sources Crossref and OpenAlex, with the in-house database of the
German Competence Network for Bibliometrics (Kompetenznetzwerk
Bibliometrie, KB), which comprises Web of Science and Scopus snapshots.
After describing the initial data sources used, the necessary
pre-processing steps to obtain eligible articles from transformative
agreements using author roles (first and corresponding) and affiliation
data are described. Overall, xxxx hybrid journals from xxx agreements
formed the basis of this study. This study covered journal articles
published between 2019 and 2023.

\subsection{Dataset characteristics}\label{dataset-characteristics}

\subsubsection{\{hoaddata\}}\label{hoaddata}

@@

@@

@@

!\textless-- -- internal -- external - matching table institutions -
journal-level data - country affiliation data --\textgreater{}

@@

\section{discussion}\label{discussion}

>>>>

% bibliography
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
  {\begin{list}{}{%
   \setlength{\itemindent}{0pt}
   \setlength{\leftmargin}{0pt}
   \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
    \setlength{\leftmargin}{\cslhangindent}
    \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
   \setlength{\itemsep}{#2\baselineskip}}}
  {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

The aim of this study is to demonstrate the suitability of open
scholarly data sources for assessing the impact of transformative
agreements on hybrid open access. To achieve this, results from
hoaddata, an openly available collection of open research information
regarding hybrid open access, was compared with the established
bibliometric databases Web of Science and Scopus. After describing the
initial data sources used, the necessary pre-processing steps to obtain
eligible articles from transformative agreements using open access
evidence, author roles (first and corresponding) and affiliation data
are presented. Overall, xxxx hybrid journals from xxx agreements that
published at least one open access article between 2019 and 2023. formed
the basis of this study.

\subsection{Data sources}\label{data-sources}

\subsubsection{hoaddata}\label{hoaddata}

hoaddata, developed and maintained by the author, is an R data package
comprising information about the uptake of hybrid open access since 2017
from several openly available data sources. It combines article-level
metadata from Crossref and OpenAlex with transformative agreement
information from the cOAlition S Journal Checker Tool (JCT), which links
journal and institutional data to agreements in the ESAC registry.

More specifically, hoaddata uses Crossref, a DOI registration agency,
for obtaining journal publication volume and open access status through
Creative Commons licence information relative to the published version
(``version of record''). Because of limited affiliation metadata in
Crossref (\url{https://doi.org/10.31222/osf.io/smxe5}), hoaddata sources
first-author affiliations from OpenAlex. While the country in which a
first-author was located was used to aggregate country-level statistics,
Research Organization Registry (ROR) identifiers (ROR-ID) were used in
conjunction with ESAC registry information on the duration of an
agreement was used to estimate whether an article was published under
active transformative agreements. This matching benefited from the
availability of the ROR-IDs in both sources. To improve the matching,
JCT data were enriched to include associated institutions, such as
unviersity hospitals.

hoaddata follows good practices for computational reproducibility using
R. The package, which includes data, code, a test suite and
documentation, is openly available on GitHub. To ensure computational
reproducibility while aggregating the data, a GitHub Actions continuous
integration and delivery (CI/CD) workflow interfaces with the SUB
Göttingen's open scholarly data warehouse based on Google BigQuery,
which provides high-performant programmatic access to monthly snapshots
of Crossref and OpenAlex. The package has been regularly updated since
2022 and the version including the computation log is available on
GitHub.

hoaddata is used as a data basis of the Hybrid Open Access Dashboard, a
data analytics services for library consortia and publishers to track
the uptake of hybrid open access through transformative agreements. It
was also used in bibliometric research (Jahn 2025).

@@

Clarivate Analytics' Web of Science (WoS) is a well-established
proprietary bibliometric database consisting of several collections
(Birkle et al., 2020). The collections considered in this study were the
Science Citation Index Expanded (SCIE), the Social Sciences Citation
Index (SSCI) and the Arts \& Humanities Citation Index (AHCI).

The WoS provides important data points for studying spending on open
access: author affiliations and roles, differentiation of journal
articles into document types representing different types of journal
contributions, such as original articles or reviews, and open access
status information derived from OurResearch's Unpaywall, the same
provider as Openalex. However, it lacks information about journals and
articles under transformative agreements.

For programmatic access to article-level data, the database of the
Kompetenznetzwerk Bibliometrie (KB) in Germany is used to access
bibliometric data. The KB processes raw XML data provided by Clarivate
Analytics, which is provided as an in-house PostgreSQL database under a
uniform schema. To support reproducibility, KB maintains annual
snapshots of the database. Accordingly, this study used the annual
snapshot from April 2024, which is considered to cover almost the entire
previous publication year (Schmidt et al., 2024).

@@

Elsevier's Scopus, launched in 2004, is another widely used proprietary
bibliometric database for measuring research (Baas et al., 2020).
Similar to the Web of Science, Scopus is selective with regard to the
journals it indexes. However, its coverage is substantially more
extensive than that of the Web of Science Core collection (Singh et al.,
2021; Visser et al., 2021). With detailed metadata about article types,
open access status information derived from Unpaywall, author roles, and
disambiguated affiliations, Scopus also contains important data to
assess open access uptake, although no direct information regarding
transformative agreements was available at the time of the study.

This study used the Scopus annual snapshot of April 2024 as provided by
the KB. The same KB curation effort was applied to the Scopus raw data
as for the Web of Science (Schmidt et al., 2024).

@@

@@

\section*{discussion}\label{discussion}
\addcontentsline{toc}{section}{discussion}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Baas_2020}
Baas, J., Schotten, M., Plume, A., Côté, G., \& Karimi, R. (2020).
Scopus as a curated, high-quality bibliometric data source for academic
research in quantitative science studies. \emph{Quantitative Science
Studies}, \emph{1}(1), 377--386.
\url{https://doi.org/10.1162/qss_a_00019}

\bibitem[\citeproctext]{ref-Birkle_2020}
Birkle, C., Pendlebury, D. A., Schnell, J., \& Adams, J. (2020). Web of
science as a data source for research on scientific and scholarly
activity. \emph{Quantitative Science Studies}, \emph{1}(1), 363--376.
\url{https://doi.org/10.1162/qss_a_00018}

\bibitem[\citeproctext]{ref-schmidt_2024_13935407}
Schmidt, M., Rimmert, C., Stephen, D., Lenke, C., Donner, P., Gärtner,
S., Taubert, N., Bausenwein, T., \& Stahlschmidt, S. (2024). \emph{The
data infrastructure of the {German Kompetenznetzwerk Bibliometrie}: An
enabling intermediary between raw data and analysis}. Zenodo.
\url{https://doi.org/10.5281/zenodo.13935407}

\bibitem[\citeproctext]{ref-Singh_2021}
Singh, V. K., Singh, P., Karmakar, M., Leta, J., \& Mayr, P. (2021). The
journal coverage of web of science, scopus and dimensions: A comparative
analysis. \emph{Scientometrics}, \emph{126}(6), 5113--5142.
\url{https://doi.org/10.1007/s11192-021-03948-5}

\bibitem[\citeproctext]{ref-Visser_2021}
Visser, M., Eck, N. J. van, \& Waltman, L. (2021). Large-scale
comparison of bibliographic data sources: {Scopus, Web of Science,
Dimensions, Crossref, and Microsoft Academic}. \emph{Quantitative
Science Studies}, \emph{2}(1), 20--41.
\url{https://doi.org/10.1162/qss_a_00112}

\end{CSLReferences}
